Bayesian regression facilitates quantitative modelling of cell metabolism

### 2023-10-31 paper/main.qmd
edits

MCMC sampling for our comparison input yielded 200 samples in 625 minutes;
Laplace sampling yielded the same number of samples in only one minute.
The diagnostics indicated that our algorithm was able to find the maximum
a posteriori parameter configuration, approximate the Hessian and use these
quantities to generate approximate posterior samples. The results can be found
at <https://github.com/biosustain/Methionine_model/tree/main/results>.

@@

@@

Importantly, this form of analysis takes into account all modelled sources of
uncertainty, including uncertainty about the true values of the flux in each
dataset, and propagates this uncertainty into the final conclusion. Our result
shows that Maud could be used in this realistic case not only to provide a user
with an explanation for an observed difference in fluxes, but also a reasonable
judgement as to the explanation's robustness.

>>>>

MCMC sampling for our comparison input yielded 800 samples in 625 minutes;
Laplace sampling yielded the same number of samples in only one minute.
The diagnostics indicated that our algorithm was able to find the maximum
a posteriori parameter configuration, approximate the Hessian and use these
quantities to generate approximate posterior samples. The results can be found
at <https://github.com/biosustain/Methionine_model/tree/main/results>.

@@

Results of removing concentration measurements for the metabolite $ahcys_c$
from our case study dataset.  (**Top**) A comparison of metabolite concentration
residuals between the full measurement dataset (blue-green) and the missing-data
dataset (grey), displayed on natural logarithmic scale. The missing-data
model was unable to estimate the withheld $ahcys_c$ concentrations.  (**Bottom
Left**) The marginal posterior distribution for the Michaelis constant
$K_m^{AHC1,ahcys_c}$ in each model, alongside the true parameter value used to
generate both datasets. The true value is recovered by the complete-data model
but not by the missing-data model.  (**Bottom Right**) The distribution of total
log-likelihood for out-of-sample flux measurements in both models. There is a
significant overlap between the two distributions, suggesting that removing the
$ahcys_c$ measurement did not cause catastrophic prediction failure. However,
overall the complete-data model tended to make better predictions.

@@

Importantly, this form of analysis takes into account all modelled sources of
uncertainty, including uncertainty about the measured values of the flux in each
dataset. Our result shows that Maud could be used in this realistic case not
only to provide a user with an explanation for an observed difference in fluxes,
but also a reasonable judgement as to the explanation's robustness.
